---
title: "Quant Cell Bio Final Project"
output: html_notebook
---

Define LOOCV
```{r}
# Leave One Out Cross Validation
loocv <- function(model, actual, predicted) {
  num = actual - predicted
  denom = 1 - hatvalues(model)
  sol = mean((num/denom)^2)
  sol
}
```

Define Kernel
```{r}
# one-dimensional smoothing Gaussian kernel, code taken from Lecture 21
kernel = function(x,y,grid,h){
  ### kernel regression estimator at a grid of values ### one dimension only
  ### return m.hat(u) for all u in grid
  n= length(x)
  k = length(grid)
  m.hat = rep(0,k)
  for(i in 1:k){
    w = dnorm(grid[i],x,h)
    m.hat[i] = sum(y*w)/sum(w)
  }
  return(m.hat)
}

# gives us the fitted values of the kernel
kernel.fitted = function(x,y,h){
  ### kernel regression
  ### fitted values and diagonal of smoothing matrix
  n = length(x)
  m.hat = rep(0,n)
  S = rep(0,n)
  for(i in 1:n){
    w = dnorm(x[i],x,h)
    w = w/sum(w)
    m.hat[i] = sum(y*w)
    S[i] = w[i]
  }
  return(list(fitted=m.hat,S=S))
  }
```




FEATURE SET ONE
read the data from the existing CSV and process it
```{r}
data = read.csv("/Users/yb/Desktop/qcb-final/featureset1.csv", sep=",", header=TRUE)
scale = max(log(data$avgarea))
dataset = data.frame(x = data$time / 1800, numsprays = data$numsprays, y = log(data$avgarea) / scale)
dataset = subset(dataset, y != -Inf & y != 0, select=c(x, y, numsprays))
```



TREATMENT: 0 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 0, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 3 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 3, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 6 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 6, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 9 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 9, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```











FEATURE SET TWO
read the data from the existing CSV
```{r}
data = read.csv("/Users/yb/Desktop/qcb-final/featureset2.csv", sep=",", header=TRUE)
scale = max(log(data$area))
dataset = data.frame(x = data$time / 1800, numsprays = data$numsprays, y = log(data$area) / scale)
dataset = subset(dataset, y != -Inf & y != 0, select=c(x, y, numsprays))
```



TREATMENT: 0 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 0, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 3 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 3, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 6 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 6, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```



TREATMENT: 9 SPRAYS
Extract data for this treatment
```{r}
groupdata = subset(dataset, numsprays == 9, select=c(x, y, numsprays))
```
Create a logistic model for the data, plot, and evaluate
```{r}
logistic_model <- glm(y ~ x, data = groupdata, family = "binomial")

log_predicted_data <- data.frame(x=seq(min(groupdata$x), max(groupdata$x), len=500))
log_predicted_data$y = predict(logistic_model, log_predicted_data, type="response")
cv = mean(((y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
log_loocv = loocv(logistic_model, groupdata$y, log_predicted_data$y)
log_loocv

plot(y ~ x, data=groupdata)
lines(y ~ x, log_predicted_data, lwd=2, col="red")
```
Create kernel models for the data, plot, and evaluate
```{r}
H = c(10, 20, 30)
cv = rep(0, length(H))

for(i in 1:length(H)) {
  h = H[i]
  grid = seq(0,300,length=100000)
  m_hat = kernel(groupdata$x, groupdata$y, grid, h)
  plot(groupdata$x, groupdata$y, col="gray", xlim=c(0, 300), ylim=c(0, 1))
  par(new=TRUE)
  plot(grid, m_hat, col="red", pch=20, cex=0.15, xlim=c(0, 300), ylim=c(0, 1), xlab="", ylab="")
  kernel_predicted_data = kernel.fitted(groupdata$x, groupdata$y, h)
  cv[i] = mean(((groupdata$y - kernel_predicted_data$fitted)/(1-kernel_predicted_data$S))^2)
}
cv
```

